MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-09-17 MAESTROeX_I/intro.tex
starting on response

\section{Introduction} \label{sec:intro}
Many astrophysical flows are highly subsonic.  In this regime,
sound waves carry sufficiently little energy that they do not
significantly affect the convective dynamics of the system.  In many
of these flows, modeling long-time convective dynamics are of
interest, and numerical approaches based on compressible hydrodynamics
are intractable, even on modern supercomputers.  One approach to this
problem is to use low Mach number models.  In a low Mach number
approach, sound waves are eliminated from the governing equations
while retaining compressibilitiy effects due to, e.g., nuclear energy
release, stratification, compositional changes, and thermal diffusion.  When the Mach
number (the ratio of the characteristic fluid velocity over the
characteristic sound speed; Ma $= U/c$) is small, the resulting system
can be numerically integrated with much larger time steps than a
compressible model.  Specifically, the time step increase is at least
a factor of $\sim 1/{\rm Ma}$ larger.  Each time step is more
computationally expensive due to the presence of additional linear
solves, but for many problems of interest the overall gains in
efficiency can easily be an order of magnitude or more.

Low Mach number models have been developed for a variety of contexts
including combustion \citep{day2000numerical}, terrestrial atmospheric
modeling \citep{durran:1989,oneill:2014,duarte2015low}, and elastic
solids \citep{abbate2017all}.  For astrophysical applications, a
number of approaches to modeling low Mach number flows have been
developed in recent years.  One of the more similar approaches to ours is
\cite{Lin:2006}, however this approach is only first-order accurate and does
not account for atmospheric expansion.  There are also semi-implicit all-Mach
number solvers, where the Euler equations are split into an acoustic
part and an advective part
\citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
The fast acoustic waves are then solved using implicit time
integration, while the slow material waves are solved explicitly.
Another approach is to use preconditioned all-Mach number solvers
\citep{Miczek2014,Barsukow2016}, where the numerical flux is
multiplied by a preconditioning matrix.  This reduces the stiffness of
the system at low Mach numbers, while retaining the correct scaling
behavior. In the reduced speed of sound technique (RSST) and related
methods, the speed of sound is artificially reduced by including a
suitable scaling factor in the continuity equation, reducing the
restriction on the size of the time step
\citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
Finally, there are fully implicit time integration codes for the
compressible Euler equations
\citep{Viallet2011,kifonidis:2012,Viallet2015,Goffrey2016}.  The MUSIC code
uses fully implicit time integration for the compressible Euler
equations, which therefore allows for arbitrarily large time steps.

>>>>

\section{Introduction} \label{sec:intro}
Many astrophysical flows are highly subsonic.  In this regime,
sound waves carry sufficiently little energy that they do not
significantly affect the convective dynamics of the system.  In many
of these flows, modeling long-time convective dynamics are of
interest, and numerical approaches based on compressible hydrodynamics
are intractable, even on modern supercomputers.  \replaced{One approach to this
problem is to use low Mach number models.  In a low Mach number
approach,}{Our approach to this problem is to use a low Mach number model where} 
sound waves are eliminated from the governing equations
while retaining compressibilitiy effects due to, e.g., nuclear energy
release, stratification, compositional changes, and thermal diffusion.
When the Mach number (the ratio of the characteristic fluid velocity over the
characteristic sound speed; Ma $= U/c$) is small, the resulting system
can be numerically integrated with much larger time steps than a
compressible model.  Specifically, the time step increase is at least
a factor of $\sim 1/{\rm Ma}$ larger.  Each time step is more
computationally expensive due to the presence of additional linear
solves, but for many problems of interest the overall gains in
efficiency can easily be an order of magnitude or more.

Low Mach number models \added{that do not contain acoustic waves}
have been developed for a variety of contexts
including combustion \citep{day2000numerical}, terrestrial atmospheric
modeling \citep{durran:1989,oneill:2014,duarte2015low}, and elastic
solids \citep{abbate2017all}.  
For astrophysical applications, a
number of approaches to modeling low Mach number flows have been
developed in recent years.  One of the more similar approaches to ours is
\cite{Lin:2006}, however this approach is only first-order accurate and does
not account for atmospheric expansion.
\added{Other approaches model acoustic wave propagation with various strategies
to efficiently and robustly handle the low Mach number limit.}
There are \deleted{also} semi-implicit all-Mach
number solvers, where the Euler equations are split into an acoustic
part and an advective part
\citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
The fast acoustic waves are then solved using implicit time
integration, while the slow material waves are solved explicitly.
Another approach is to use preconditioned all-Mach number solvers
\citep{Miczek2014,Barsukow2016}, where the numerical flux is
multiplied by a preconditioning matrix.  This reduces the stiffness of
the system at low Mach numbers, while retaining the correct scaling
behavior \added{in an acoustic limit of the Euler equations that 
retains $\mathcal{O}$(Ma) pressure fluctuations}.
In the reduced speed of sound technique (RSST) and related
methods, the speed of sound is artificially reduced by including a
suitable scaling factor in the continuity equation, reducing the
restriction on the size of the time step
\citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
Finally, there are fully implicit time integration codes for the
compressible Euler equations
\citep{kifonidis:2012,Viallet2015,Goffrey2016}.
\explain{removed outdated Viallet2011 citation and corrected the typo in the Viallet2015 citation}
The MUSIC code
uses fully implicit time integration for the compressible Euler
equations, which therefore allows for arbitrarily large time steps.
